### 2025-04-15 resources.qmd

---
title: "R implementations"
---

### IDA report for regression models

[Bacteremia data set](https://zenodo.org/records/7554815) by Georg Heinze

<https://stratosida.github.io/regression-regrets/>

### IDA report for longitudinal studies

Grip strength data set from the [SHARE project](https://share-eric.eu)

<https://stratosida.github.io/longitudinal/>