### 2025-04-16 index.qmd
Update index.qmd with TG2 content

---
title: "Selection of Variables and Functional Forms in Multivariable Analysis"
---

## Aims

Derive guidance for variable and function selection  in multivariable analysis.

## Focus

The main focus of TG2 is to identify influential variables and gain insight into their individual and joint relationship with the outcome. Two of the (interrelated) main challenges are selection of variables for inclusion in a multivariable explanatory model, and choice of the functional forms for continuous variables

## STRATOS Initiative

Topic group 2 is a member of the [STRATOS](https://stratos-initiative.org) Initiative (STRengthening Analytical Thinking for Observational Studies) which is a large collaboration of experts in many different areas of biostatistical research. Ongoing research, discussions and activities within STRATOS are conducted in nine [topic groups](https://stratos-initiative.org/en/groups) and several cross-cutting [panels](https://stratos-initiative.org/en/panels).

![2nd General Meeting of the STRATOS members at the Banff International Research Station (BIRS), Banff, Canada.](assets/STRATOS_Banff_second_meeting.png){fig-align="center"}