### 2025-04-16 members.qmd
Update members.qmd with TG2 content

---
title: "Members"
---

[Georg Heinze](https://www.meduniwien.ac.at/researcher/georg_heinze) (chair): Institute of Clinical Biometrics, Center for Medical Statistics, Informatics and Intelligent Systems, Medical University of Vienna,  Austria.

**Aris Perperoglou (chair)**:  Data Science & AI,   AstraZeneca R&D Biopharmaceuticals , Cambridge UK

**Willi Sauerbrei (chair)**: Institute for Medical Biometry and Medical Informatics, Medical Center - University of Freiburg, Germany

**Michal Abrahamowicz**: Faculty of Medicine, McGill University, Canada

**Heiko Becher**: Institute for Medical Biometry and Epidemiology, University Medical Center Hamburg-Eppendorf, Hamburg, Germany

**Harald Binder**:  Institute for Medical Biometry and Medical Informatics, Medical Center - University of Freiburg, Germany

**Daniela Dunkler**: Institute of Clinical Biometrics, Center for Medical Statistics, Informatics and Intelligent Systems, Medical University of Vienna,  Austria

**Rolf Groenwold**:  Leiden University Medical Centre, Leiden, The Netherlands

**Frank Harrell**: School of Medicine, Vanderbilt University Medical Center, USA

**Nadja Klein**: School of Business and Economics , Humboldt-University of Berlin, Germany

**Geraldine Rauch**:  Institute of Biometry and Clinical  Epidemiology, Charité – Universitätsmedizin Berlin , Germany

**Patrick Royston**: Department of Statistical Science, UCL, UK

**Matthias Schmid**:  Institute for Medicine Biometry, Informatics and Epidemiology, Bonn, Germany

## Early career adjunct members

**Michael Kammer**: Institute of Clinical Biometrics, Center for Medical Statistics, Informatics and Intelligent Systems, Medical University of Vienna,  Austria

**Edwin  Kipruto**:  Institute for Medical Biometry and Medical Informatics, Medical Center - University of Freiburg, Germany

**Kim Luijken**:  Leiden University Medical Centre, Leiden, The Netherlands

**Christine Wallisch**: Institute of Clinical Biometrics, Center for Medical Statistics, Informatics and Intelligent Systems, Medical University of Vienna,  Austria