### 2025-04-16 papers.qmd
Test papers.qmd collapsible update

---
title: "TG2 Publications"
---

<details>
  <summary>Click for Answer</summary>
    ```{r, echo=FALSE, eval=TRUE}
    hist(1:10)
    ```
</details>