### 2025-05-02 about.qmd
Full update

---
title: "About this website"
---

**Andreas Klinger**: set-up and maintenance

**Georg Heinze**: supervision and revision

Last changes: 2 May 2025.

### Acknowledgements

Many thanks to Marianne Huebner, who designed the website of the STRATOS Topic Group 3, which the website of Topic Group 2 is based on.

### Address for correspondence

STRATOS TG2 c/o Institute of Clinical Biometrics, center for Medical Data Science, Medical University of Vienna, Vienna, Austria. E-mail: stratos.tg2@gmail.com.