### 2025-05-02 members.qmd

---
title: "Members"
---

**[Georg Heinze](https://www.meduniwien.ac.at/researcher/georg_heinze) (chair)**: Institute of Clinical Biometrics, Center for Medical Data Science, Medical University of Vienna,  Austria.

**Aris Perperoglou (chair)**:  Data Science & AI,   AstraZeneca R&D Biopharmaceuticals , Cambridge UK

**Willi Sauerbrei (chair)**: Institute for Medical Biometry and Medical Informatics, Medical Center - University of Freiburg, Germany

**Michal Abrahamowicz**: Faculty of Medicine, McGill University, Canada

**Harald Binder**:  Institute for Medical Biometry and Medical Informatics, Medical Center - University of Freiburg, Germany

**Daniela Dunkler**: Institute of Clinical Biometrics, Center for Medical Data Science, Medical University of Vienna,  Austria

**Frank Harrell**: School of Medicine, Vanderbilt University Medical Center, Nashville, TN, USA

**Marc Henrion**: Malawi Liverpool Wellcome Research Programme, Blantyre, Malawi and Department of Clinical Sciences, Liverpool School of Tropical Medicine, Liverpool, UK

**Michael Kammer**: Division of Nephrology, Department of Medicine III and Institute of Clinical Biometrics, Center for Medical Data Science, Medical University of Vienna,  Austria

**Mariana Nold**: Department of Sociology, Friedrich-Schiller University of Jena, Jena, Germany

**Matthias Schmid**:  Institute for Medicine Biometry, Informatics and Epidemiology, University of Bonn, Bonn, Germany

## Early career adjunct members

**Gregor Buch**: Preventive Cardiology and Preventive Medicine, Department of Cardiology, University Medical Center of the Johannes Gutenberg University Mainz, Mainz, Germany

**Edwin  Kipruto**:  Institute for Medical Biometry and Medical Informatics, Medical Center - University of Freiburg, Germany

**Theresa Ullmann**: Institute of Clinical Biometrics, Center for Medical Data Science, Medical University of Vienna,  Austria