### 2025-05-02 about.qmd
small changes

---
title: "About this website"
---

**Andreas Klinger**: set-up and maintenance

**Georg Heinze**: supervision and revision

Last changes: 2 May 2025.

### Acknowledgements

Many thanks to Marianne Huebner, who designed the website of the STRATOS Topic Group 3, which the website of Topic Group 2 is based on.

### Address for correspondence

STRATOS TG2

c/o Institute of Clinical Biometrics,

Center for Medical Data Science,

Medical University of Vienna,

Spitalgasse 23, 1090 Vienna, Austria.

E-mail: stratos.tg2@gmail.com.